\documentclass[handout]{ximera}
\input{preamble}
\title{Linear systems of equations}

\begin{document}
\begin{abstract}
  Linear algebra is the study of linear functions.
\end{abstract}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A {\it linear function in one variable} is of the form $f(x) = ax+b$ where $x$ is a variable and $a,b$ are numbers (or scalars, as they are referred to in linear algebra). It is {\it homogeneous} if $b=0$. Similarly, a {\it linear function in n variables} is one of the form
\[
f(x_1,x_2,\dots,x_n) = a_1x_1 + a_2x_2 + \dots a_nx_n + b
\]
where the $x_i$ are variables (or unknowns) and the $a_i$ are scalars. Again, the linear function is called {\it homogeneous} if the constant term $b$ is zero. Next, a {\it linear equation in n variables} is one of the form
\[
a_1x_1 + a_2x_2 + \dots a_nx_n =  b
\]
where the expression on the left is a linear homogeneous function in $n$ variables, and the term on the right is a constant. So, for example, 
\begin{equation}\label{eqn:ex1}
3x_1 - 7x_2 + 11x_3 = 14
\end{equation}

is a linear equation in three variables - $x_1, x_2$, and $x_3$. A {\it solution} to a linear equation is an assignment of values to each of the variables appearing which makes the equation hold true. Returning to the example in the equation (\ref{eqn:ex1}), we see
\[]
x_1 = 1, x_2 = 0, x_3 = 1
\]
is a solution, because when substituted into the left-hnd side it results in the value 14. Finally, a {\it system of equations} is a collection of linear equations in the same set of variables, or unknowns
\begin{align*}
a_{11}x_1 + a_{12}x_2  + {}\cdots{}  + a_{1n}x_n &=  b_1 \\
a_{21}x_1 +  a_{22}x_2 +   {}\cdots{}  + a_{2n}x_n &= b_2 \\
\vdots \\
a_{m1}x_1 + a_{m2}x_2  + {}\cdots{} + a_{mn}x_n &=  b_m 
\end{align*}

and a {\it solution} to that system is an assignment of values to the variables which make {\it each equation} hold true. Note that systems of equations, or even a single equation, need to have a solution. To illustrate, consider the equation
\[
0x_1 + 0x_2 + 0x_3 = 4
\]
Obviously, any set of values substituted into the left-hand side of this equation will produce the value zero, which is not equal to 4. But even non-zero systems might not have a solution. As an example, consider
\begin{alignat*}{3}
x_1 && - 2x_2 && =& \phantom{1}7 \\
2x_1 && - 4x_2 && =& 16 
\end{alignat*}
The left-hand side of the second equation is twice that of the first. So if we take any solution of the first equation and plug in those values on the left-hand side of the second equation, we will always get $2*7 = 14\ne 16$. Therefore, these two equations will never be simultaneously satisfied, and so the system doesn't have a solution.
\vskip,2in
A system which has at least one solution is called {\it consistent}. If it doesn't have any solutons it is called {\it inconsistent}. The main questions we need to answer are:
\vskip.1in

\begin{question} Given a system of equations, does it have one or more solutions (in other words is it consistent)?
\end{question}
\vskip.05in
\begin{question} If it is consistent, what are the solutions?
\end{question}
\vskip.3in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
